\chapter{Canon of Repentance to our Lord Jesus Christ}

\textbf{Song 1. Tone 6.}

\textbf{Eirmos:} When Israel walked on foot in the sea as on dry land,
on seeing their pursuer Pharaoh drowned, they cried: Let us sing to God
a song of victory. Have mercy on me, O God, have mercy on me.

\textbf{Troparia:}

Now I, a burdened sinner, approach Thee, my Lord and God. But I dare not
raise my eyes to Heaven. I only pray, saying: Give me, O Lord, the sense
to weep bitterly over my deeds.

Have mercy on me, O God, have mercy on me.

O woe is me, a sinner! Wretched am I above all men. There is no
penitence in me. Give me, O Lord, tears to weep bitterly over my deeds.

Glory to the Father, and to the Son, and to the Holy Spirit:

Foolish, wretched man, you are wasting your time in idleness! Think of
your life and turn to the Lord God, and weep bitterly over your deeds.

Now and ever, and to the ages of ages. Amen.

\textbf{Theotokion:} Most pure Mother of God, look upon me, a sinner,
and deliver me from the snares of the devil, and guide me to the way of
repentance, that I may weep bitterly over my deeds.

\textbf{Song 3.}

\textbf{Eirmos:} There is none holy as Thou, O Lord my God, Who hast
exalted the power of Thy faithful, O Good One, and strengthened us on
the rock of Thy confession.

Have mercy on me, O God, have mercy on me.

\textbf{Troparia:}

When the thrones are set at the dread judgment, then the deeds of all
men will be exposed. Then alas for the sinners sent to torment! And
knowing that, my soul, repent of your evil deeds.

Have mercy on me, O God, have mercy on me.

The righteous will rejoice, but the sinners will weep. Then no one will
be able to help us, but our deeds will condemn us. Therefore, before the
end dawns, repent of your evil deeds.

Glory to the Father, and to the Son, and to the Holy Spirit:

Woe is me, a great sinner, who have defiled myself by my deeds and
thoughts. Not a tear-drop do I have, because of my hard-heartedness. Now
raise yourself from the earth, my soul, and repent of your evil deeds.

Now and ever, and to the ages of ages. Amen.

\textbf{Theotokion:} Lo, thy Son calls, O Lady, and teaches us what is
good. But, sinful as I am, I always flee from the good. But do thou, O
merciful one, have mercy on me, that I may repent of my evil deeds.

Lord, have mercy. (Thrice)

\textbf{Sedalion, tone 6:}

I think of the awful day and weep over my evil deeds. How shall I answer
the Immortal King? How shall I, a prodigal, dare to look at the Judge? O
gracious Father, Only-Begotten Son, and Holy Spirit, have mercy on me.

Glory to the Father, and to the Son, and to the Holy Spirit, now and
ever, and to the ages of ages. Amen.

\textbf{Theotokion:} Bound now with many chains of sins, and inhibited
by cruel passions, I have recourse to thee, my salvation, and cry: Help
me, O Virgin, Mother of God.

\textbf{Song 4.}

\textbf{Eirmos:} Christ is my power, my God and my Lord, the holy Church
divinely sings, crying with a pure mind, keeping festival in the Lord.

Have mercy on me, O God, have mercy on me.

\textbf{Troparia:}

Broad is the way here and conducive to indulging in pleasures, but how
bitter it will be on the last day when the soul is separated from the
body! Beware of pleasures, man, for the sake of the Kingdom of God.

Have mercy on me, O God, have mercy on me.

Why do you wrong the poor? Why do you withhold the wage of the hired
servant? Why do you not love your brother? Why do you run after pride
and lust? And so, stop these things, my soul, and repent for the sake of
the Kingdom of God.

Glory to the Father, and to the Son, and to the Holy Spirit.

O thoughtless man! How long will you busy yourself like a bee,
accumulating your fortune? For it will perish like dust and ashes soon.
But seek rather the Kingdom of God.

Now and ever, and to the ages of ages. Amen.

\textbf{Theotokion:} O Lady, Mother of God, have mercy on me, a sinner.
Strengthen me in the virtues and keep me safe, lest sudden death take me
by surprise. And bring me, O Virgin, to the Kingdom of God.

\textbf{Song 5.}

\textbf{Eirmos:} Illumine with Thy divine light, I pray, O Good One, the
souls of those who with love rise early to pray to Thee, that they may
know Thee, O Word of God, as the true God, Who recalls us from the
darkness of sin.

Have mercy on me, O God, have mercy on me.

\textbf{Troparia:}

Remember, wretched man, how you are enslaved to lies, calumnies, theft,
infirmities, wild beasts and fears, on account of your sins. O my sinful
soul, is that what you have desired?

Have mercy on me, O God, have mercy on me.

My members tremble, for with all of them I have sinned: with my eyes in
looking, with my ears in hearing, with my tongue in speaking evil, and
by surrendering the whole of myself to hell. O my sinful soul, is that
what you have desired?

Glory to the Father, and to the Son, and to the Holy Spirit.

Thou didst receive the prodigal and the robber who repented, O Saviour,
and I alone have succumbed to sinful sloth and have become enslaved to
evil deeds. O my sinful soul, is this what you have desired?

Now and ever, and to the ages of ages. Amen.

\textbf{Theotokion:} Wonderful and speedy helper of all men, help me.
Mother of God, unworthy as I am, for my sinful soul desires that.

\textbf{Song 6.}

\textbf{Eirmos:} Beholding the sea of life surging the flood of
temptations, I run to calm haven, and cry to Thee: Raise up my life from
corruption, O Most Merciful One.

Have mercy on me, O God, have mercy on me.

\textbf{Troparia:}

I have lived my life wantonly on earth and have delivered my soul to
darkness. But now I implore Thee, O merciful Lord, free me from this
work of the enemy and give me the knowledge to do Thy will.

Have mercy on me, O God, have mercy on me.

Who does such things as I do? For just like a swine lying in the mud, so
I serve sin. But do Thou, O Lord, pull me out of this vileness and give
me the heart to do Thy commandments.

Glory to the Father, and to the Son, and to the Holy Spirit:

Rise, wretched man, to God and, remembering your sins, fall down before
your Creator, weeping and groaning, for He is merciful and will grant
you to know His will.

Now and ever, and to the ages of ages. Amen.

Virgin Mother of God, protect me from evil visible and invisible, O
immaculate one, and accept my prayers and convey them to thy Son, that
He may grant me the sense to do His will.

Lord, have mercy. (Thrice).

Glory to the Father, and to the Son, and to the Holy Spirit, now and
ever, and to the ages of ages. Amen.

\textbf{Kontakion:} O my soul, why do you become rich in sins? Why do
you do the will of the devil? On what do you set your hope? Stop these
things and turn to God with tears, and cry: O Merciful Lord, have mercy
on me, a sinner.

\textbf{Oekos:} Think, my soul, of the bitter hour of death and the
awful judgment of your God and Creator. For terrible Angels will seize
you, my soul, and will lead you into the eternal fire. And so, before
your death, repent and cry: O Lord, have mercy on me, a sinner.

\textbf{Song 7.}

\textbf{Eirmos:} An Angel made the furnace throw dew on the holy
Children. But the command of God consumed the Chaldeans and prevailed
upon the tyrant to cry: Blessed art Thou, O God of our fathers.

Have mercy on me, O God, have mercy on me.

\textbf{Troparia:}

Do not hope, my soul, for corruptible wealth, and for what is unjustly
collected. For you do not know to whom you will leave it all. But cry: O
Christ our God, have mercy on me, who am unworthy.

Have mercy on me, O God, have mercy on me.

Do not trust, my soul, in your physical health, and in your
quickly-passing beauty. For you see that the strong and the young die.
But say: O Christ our God, have mercy on me, who am unworthy.

Glory to the Father, and to the Son, and to the Holy Spirit:

Remember, my soul, eternal life and the Heavenly Kingdom prepared for
the saints, and the outer darkness and the wrath of God for the evil,
and cry: O Christ our God, have mercy on me, who am unworthy.

Now and ever, and to the ages of ages. Amen.

\textbf{Theotokion:} Fall down, my soul, before the Mother of God, and
pray to her; for she is quick to help those who repent. She prays to
Christ, her Son and our God, and has mercy on me who am unworthy.

\textbf{Song 8.}

\textbf{Eirmos:} Thou didst make flame sprinkle the Saints with dew, and
didst burn the sacrifice of a righteous man with water. For Thou alone,
O Christ, dost do all as Thou willest. Thee we exalt throughout all
ages.

Have mercy on me, O God, have mercy on me.

\textbf{Troparia:}

How can I not weep when I think of death? For I have seen my brother
lying in his coffin, inglorious and hideous. What, then, do I expect?
And what do I hope for? Only grant me, O Lord, repentance before my end.
(2)

Glory to the Father, and to the Son, and to the Holy Spirit.

I believe that Thou wilt come to judge the living and the dead, and all
will stand in order, old and young, lords and princes, priests and
virgins. Where shall I find myself? Therefore I cry: grant me, O Lord,
repentance before my end.

Now and ever, and to the ages of ages. Amen.

\textbf{Theotokion:} Most pure Mother of God, accept my unworthy prayer
and preserve me from sudden death; and grant me repentance before my
end.

\textbf{Song 9.}

\textbf{Eirmos:} It is impossible for men to see God, upon Whom the
orders of Angels dare not gaze. But through thee, O all-pure one, did
the Word incarnate appear to men, and with the Heavenly Hosts we magnify
Him, and thee we call blessed.

Have mercy on me, O God, have mercy on me.

\textbf{Troparia:}

I now have recourse to you, holy Angels, Archangels, and all the
Heavenly Hosts who stand at the throne of God: pray to your Creator that
He may save my soul from eternal torment.

Have mercy on me, O God, have mercy on me.

Now I turn to you with tears, holy patriarchs, kings and prophets,
apostles and holy prelates, and all Christ's elect: help me at the
judgment, that He may save my soul from the power of the enemy.

Glory to the Father, and to the Son, and to the Holy Spirit.

Now I lift my hands to you, holy martyrs, hermits, virgins, righteous
and all the saints, who pray to the Lord for the whole world, that He
may have mercy on me at the hour of my death.

Now and ever, and to the ages of ages. Amen.

\textbf{Theotokion:} O Mother of God, help me who have strong hope in
thee; implore thy Son that He may place me on His right hand, unworthy
as I am, when He sits to judge the living and the dead. Amen.

\textbf{Prayer after the Canon}

Lord Christ our God, Who hast healed my passions through Thy Passion,
and hast cured my wounds through Thy wounds, grant me who have sinned
greatly against Thee tears of compunction. Transform my body with the
fragrance of Thy life-giving Body, and sweeten my soul with Thy Holy
Blood from the bitterness with which the foe has fed me. Lift up my
downward looking mind to Thee, and take it out of the pit of perdition,
for I have no repentance, I have no compunction, I have no consoling
tears, which uplift children to their heritage. My mind has been
darkened through earthly passions, I cannot look up to Thee in pain. I
cannot warm myself with tears of love for Thee. But, O Lord Jesus
Christ, Treasury of good gifts, give me thorough repentance and a
diligent heart to seek Thee; grant me Thy grace, and renew in me the
lineaments of Thy image. I have forsaken Thee - do not forsake me! Come
out to seek me; lead me up to Thy pasturage and number me among the
lambs of Thy chosen flock. Nourish me with them on the grass of Thy Holy
Mysteries, by the prayers of Thy Most Pure Mother and all Thy saints.
Amen.
