\chapter{Supplicatory Canon to our Sweetest Lord Jesus Christ}

A poem by Theoctistus, a Monk of Stoudion

Second Tone

ODE ONE Heirmos

In the deep of old, the Power supreme to all weaponry overwhelmed
Pharaoh's whole army. And when the Word became incarnate, He destroyed
exceedingly pernicious sin; for the glorified Lord has gloriously been
glorified.

O Sweetest Jesus, save me.

Sweetest Jesus Christ, long-suffering Jesus, heal the wounds of my soul,
O Jesus, and make sweet my heart, O most Merciful One, I pray to You, O
Jesus my Saviour, that, as one saved, I may magnify You.

O Sweetest Jesus, save me.

Sweetest Jesus Christ, O Jesus, open to me the portals of repentance, O
Jesus, You the Friend of man, and accept me, O Jesus my Saviour, as I
fall down before You and fervently implore the forgiveness of my sins.

Glory to the Father and the Son and the Holy Spirit.

Sweetest Jesus Christ, O Jesus, snatch me from the hand of deceitful
Belial, O Jesus, and make me stand at the right of You glory, O Jesus my
Saviour, delivering me from the lot of those on the left.

Both now and forever and to the ages of ages. Amen.

Theotokion

O Lady, who has given birth to Jesus our God, pray for us worthless
slaves, that by your intercessions, O immaculate one, we who are defiled
may be delivered from torment, O spotless one, and enjoy everlasting
glory.

ODE THREE

Heirmos

Establishing me on the rock of faith, You have enlarged my mouth over my
enemies; for my spirit rejoices in chanting: There is none holy as our
God, and none righteous save You, O Lord.

O Sweetest Jesus, save me.

Hearken, O my man-befriending Jesus, unto You servant calling with
compunction; and deliver me, O Jesus, from condemnation and punishment,
O only long-suffering, sweetest and most merciful Jesus.

O Sweetest Jesus, save me.

Receive You servant, O my Jesus, who falls prostrate with tears, my
Jesus, and save me as one repentant, O my Jesus, delivering me from
Gehenna, O Master, sweetest and most merciful Jesus.

Glory to the Father and the Son and the Holy Spirit.

The time which You gave me, O my Jesus, have I squandered in passions,
my Jesus. Cast me not away, O my Jesus, but call me back again, I pray,
O Master, O sweetest Jesus, and save me.

Both now and forever and to the ages of ages. Amen.

Theotokion

O Virgin, who has given birth to my Jesus, implore Him to `rescue me
from Gehenna. You alone are the protectress of the afflicted, O you who
are full of divine grace. And make me worthy of the life that ages not.
O all-blameless one.

SESSIONAL HYMN

First Tone. Thy tomb, O Saviour

O Jesus my Saviour, Who has saved the prodigal, O Jesus my Saviour, Who
has accepted the harlot: Have mercy on me also, O most Merciful Jesus;
have compassion and save me, O Jesus my Benefactor, even as You had
compassion on Manasses, O my Jesus, the only Friend of man.

ODE FOUR

Heirmos

From a Virgin did You come, not as an ambassador, nor as an angel, but
the very Lord Himself incarnate, and saved me, the whole man. Therefore,
I cry to You: Glory to Your power, O Lord.

O Sweetest Jesus, save me.

Heal. O my Jesus, the wounds of my soul, my Jesus, I pray, and snatch me
from the hand of soul-corrupting Belial, O my compassionate Jesus, and
save me.

O Sweetest Jesus, save me.

I have sinned, O my sweetest Jesus; my compassionate Jesus, save me who
flee to Your protection. O long-suffering Jesus, and make me worthy of
Your Kingdom.

Glory to the Father and the Son and the Holy Spirit.

No one, O my Jesus, has sinned as have I, the wretched one; but now I
fall prostrate and pray: Save me, O my Jesus, and bequeath me life, O my
Jesus.

Both now and forever and to the ages of ages. Amen. Theotokion

O all-hymned one, who has given birth to the Lord Jesus, implore Him to
deliver from torment all who hymn you and rightly call you the Mother of
God.

ODE FIVE

Heirmos

O Christ my Saviour, You are the Light of those that lie in darkness and
Salvation of the despairing, unto You do I rise early, O King of peace.
Enlighten me with Your illumination, for besides You I know none other
God.

O Sweetest Jesus, save me.

You are the Light of my mind, O my Jesus; You are the Salvation of my
despairing soul, O Saviour. O my Jesus, deliver me from torment and
Gehenna, as I cry: Save me, the wretched one, O Christ my Jesus.

O Sweetest Jesus, save me.

Utterly cast down to shameful passions, O my Jesus, I now cry: Stretch
down to me a helping hand, O my Jesus, and pluck me out as I cry: Save
me, the wretched one, O Christ my Jesus.

Glory to the Father and the Son and the Holy Spirit.

Carrying about a mind defiled, I call to You, O Jesus: Cleanse me of the
defilement of sin, and redeem me who have slipped down into the depths
of evil through ignorance, and save me, O Saviour, my Jesus, I pray.

Both now and forever and to the ages of ages. Amen. Theotokion

O Maiden Mother of God, who has given birth to Jesus, implore Him to
save all Orthodox monastics and laity, and to deliver from Gehenna those
who cry: We know no certain protection save yours.

ODE SIX

Heirmos

Encompassed round about by an abyss of offences, I appeal to the
unfathomable abyss of Your compassion: Raise me up from corruption, O
God.

O Sweetest Jesus, save me.

O most merciful Christ, my Jesus, accept me who confess my sins, O
Master, and save me, O Jesus, and snatch me from corruption, O Jesus.

O Sweetest Jesus, save me.

O my Jesus, no one else has been so prodigal as I, the wretched one, O
man-befriending Jesus; but do save me, O Jesus.

Glory to the Father and the Son and the Holy Spirit.

O my Jesus, with my passions I have surpassed the harlot and the
prodigal, Manasses and the publican, O my Jesus, and the thief and the
Ninevites, O Jesus.

Both now and forever and to the ages of ages. Amen. Theotokion

O you who did conceive my Jesus Christ, O only undefiled and immaculate
Virgin: even now, by the hyssop of your intercessions, cleanse me, the
defiled one.

KONTAKION

Fourth Tone

O Sweetest Jesus, Light of the world, enlighten the eyes of my soul, O
Son of God, with Your divinely radiant splendour, that I may praise You,
the unwaning Light.

When the golden image was adored in the plain of Deira, Your three
Children trampled underfoot the ungodly decree; and though cast into the
midst of the fire, as ones refreshed they chanted: Blessed are You, the
God of our Fathers.

O Sweetest Jesus, save me.

O Christ Jesus, no one on earth has ever sinned, O my Jesus, as have I,
the wretched one and prodigal. Therefore, I cry to You, my Jesus, have
compassion on me as I sing: Blessed are You, the God of our Fathers.

O Sweetest Jesus, save me.

O Christ Jesus, I cry: Nail me down with the fear of You, O my Jesus,
and guide me to Your calm haven now, O my compassionate Jesus, that as
one saved I may sing to You: Blessed art You, the God of our Fathers.

Glory to the Father and the Son and the Holy Spirit.

O Christ Jesus, ten thousand times have I, the hapless one, promised You
repentance, O my Jesus, but, wretch that I am, I lied to You. Therefore,
I cry to You, my Jesus: Enlighten my soul which yet remains unfeeling, O
Christ, the God of our Fathers.

Both now and forever and to the ages of ages. Amen.

Theotokion

O you who gave birth to Jesus Christ awesomely and supernaturally, O
all-blameless one, implore Him, O Maiden, to forgive me all my sins
which I have committed against my human nature, that as one saved I may
cry: Blessed are you who conceived God in the flesh.

O you works, praise you the Lord God Who descended into the fiery
furnace with the Hebrew Children and changed the flame into dew; and
supremely exalt Him unto all the ages.

O Sweetest Jesus, save me.

I implore You, O my Jesus: As You redeemed the harlot from many sins, O
my Jesus, likewise redeem me, O Christ my Jesus, and cleanse my foul
soul, O my Jesus.

O Sweetest Jesus, save me.

O Jesus, having yielded to irrational pleasures. I have become
irrational, O my Jesus; and wretch that I am, I have truly and pitifully
become like the beasts, O Saviour. Therefore, O Jesus, deliver me from
irrationality.

Glory to the Father and the Son and the Holy Spirit.

Having fallen, O Jesus, into the hands of soul-corrupting thieves, I
have been stripped now of my divinely-woven garment, O Jesus, and
prostrate I lie all bruised with wounds, O my Christ, do pour on me oil
and wine.

Both now and forever and to the ages of ages. Amen.

Theotokion

O Theotokos Mary, who inexplicably held the Christ, my Jesus and God:
Implore Him to save from perils your servants and them that praise you,
O Virgin who knew no wedlock.

With one accord, let us the faithful magnify with hymns God the Word,
Who came forth from God, and Who was inexpressibly incarnate of the holy
Virgin for our sakes, and Who in His ineffable wisdom came to renew Adam
who fell grievously when he partook of corruption.

O Sweetest Jesus, save me.

I have surpassed, O my Jesus, Manasses and the publican, the harlot and
the prodigal, O compassionate Jesus, and the thief, O my Jesus, through
all my shameful and unbecoming deeds, O Jesus; but do forestall me, O my
Jesus, and save me.

O Sweetest Jesus, save me.

By my passions, O my Jesus, pitifully have I, the wretched one,
surpassed all those from Adam who have sinned, both those before the Law
and in the Law, O Jesus, and after the Law and Grace, O my Jesus; but by
Your judgments save me, O my Jesus.

Glory to the Father and the Son and the Holy Spirit.

May I not be parted from Your ineffable glory, my Jesus, nor may the
portion on the left, O Jesus, fall to me, O sweetest Jesus; but set me
on the right hand with Your sheep and give me rest, O Christ my Jesus,
since You are compassionate.

Both now and forever and to the ages of ages. Amen.

Theotokion

O Theotokos, who alone held Jesus, O Virgin Mary who knew not wedlock: O
pure one, implore Him, your Son and Creator, to be gracious to us, that
we who have recourse to you may be delivered from temptations and
perils, and from the fire that is to come.

Stichera (a poem of Theoctistus): Plagal of Second Tone. Having placed
their every hope

O my sweetest Jesus, You the joy and Gladness of my soul, Jesus, You the
cleansing of my mind, O my Master and Lord most merciful: Jesus, save
me, O Jesus my Saviour; O my Jesus, You Almighty One, do not abandon me;
O my Saviour Jesus, have mercy and deliver me from punishment, O my
Jesus, and deem me worthy of the lot of the saved, O my Jesus, and
number me with the choir of all Your blessed and chosen ones, O Jesus,
Friend of man.

O my sweetest Jesus, You the glory of the Apostles, O my Jesus, You the
boast of the Martyrs, Master Who are omnipotent: Jesus, do save me, O
Jesus my Saviour; my Jesus, You the most beautiful One, have mercy on
me, O Saviour Jesus, who runs to You, by the entreaties of her that
carried You and, O Jesus, of all Your Saints, and of all the Prophets, O
Saviour Jesus, and give the Paradise of delight to me, O Jesus, Friend
of man.

Glory to the Father and the Son and the Holy Spirit.

O my sweetest Jesus, You the glory of the monastics, O long-suffering
Jesus, You the ascetics' blest adornment and great delight: Jesus, save
me, O Jesus my Saviour; my Jesus You are supremely good, snatch me out
of the hand of the dragon, O Saviour Jesus Christ, and free me from his
snares, O my Saviour Jesus, and now lead me up, my Saviour, out of the
lowest pit, O Jesus, and number me, O Jesus, with Your sheep at Your
holy, blessed right hand.

Both now and forever and to the ages of ages. Amen.

Theotokion

O entrust me not, I pray, to any human protection, O our Lady, holy one,
but accept the prayer of your supplicant. Sorrow has fettered me, and I
am unable to endure and bear the demons' darts; a shelter have I not,
neither place to run, I, the wretched one; embattled from all sides am
I, and no consolation have I but you, Mistress of creation, protection
and hope of faithful ones: Turn not away when I pray to you; do that
which will profit me.

A prayer to our Lord Jesus Christ

O Master Christ God, Who by Your Passion has healed my passions, and by
You wounds has healed my wounds: grant tears of compunction unto me who
have greatly offended; temper my body with the fragrance of Your
life-creating Body, and sweeten my soul with Your precious Blood, taking
away the bitterness which the contending adversary has made me drink.
Lift my mind up to You, since it has been drawn down below, and lead me
up from the pit of destruction; for I have no repentance, I have no
compunction, I have not a tear of supplication as children which shall
lead me up unto my inheritance. I have been darkened in mind by the
passions of life and I am not able to look up to You in my travail; I
cannot rouse myself to fervour with tears of love for You. But, O
Master, Lord Jesus Christ, Treasury of good things, grant me complete
repentance, and a suffering heart that shall seek You out. Grant Your
grace to me and renew in me the likeness of Your image. I have abandoned
You; do not abandon me. Come forth and seek me out, lead me into Your
pasture, number me together with the sheep of Your elect flock, and
nourish me together with them on the verdure of Your divine Mysteries,
by the intercessions of Your all-pure Mother and of all Your Saints.
Amen.
