\chapter{Moleben}

\emph{When done alone all parts are read. Nothing is skipped. All parts may be read plainly.}

\begin{center}
\rule{0.5\linewidth}{0.5pt}
\end{center}

\textbf{\emph{Senior Reader:}}
 Through the prayers of our holy Fathers, O Lord Jesus Christ our God,~have~mercy on us.

\textbf{\emph{Reader:}}
~Amen.~Glory to Thee, our God, glory to Thee.

O Heavenly King, Comforter,~Spirit~of Truth, Who art everywhere present and fillest all things, Treasury of good things and Giver of life: Come and dwell in us, and cleanse us of all impurity, and save our souls, O Good One.

Holy God, Holy Mighty, Holy Immortal, have mercy on
us.
~\textbf{\emph{(Thrice)}}

Glory to the Father, and to the Son, and to the Holy Spirit, both now and ever, and unto the ages of ages.~Amen.

O Most Holy Trinity, have mercy on us. O Lord, blot out our sins. O Master, pardon our iniquities. O Holy One, visit and heal our infirmities for Thy name's sake.

Lord, have mercy.
~\textbf{\emph{(Thrice)}}

Glory to the Father and to the Son and to the Holy Spirit, both now and ever, and unto the ages of ages.~Amen.

Our Father, Who art in the Heavens, hallowed be Thy Name. Thy Kingdom come, Thy will be done, on earth as it is in Heaven. Give us this day our daily bread, and forgive us our debts, as we forgive our debtors; and lead us not into temptation, but deliver us from the evil one.

\textbf{\emph{Senior~Reader:}}
 O Lord, Jesus Christ, Son of God, have mercy on us.

\textbf{\emph{Reader:}}
~Amen.

Lord, have mercy.
~\textbf{\emph{(12 times)}}

Glory to the Father and to the Son and to the Holy Spirit, both now and ever, and unto the ages of ages.~Amen.

O come, let us worship God our King.

O come, let us worship and fall down before Christ our King and God.

O come, let us worship and fall down before Christ Himself, our King and God.

\subsubsection{Psalm 142}\label{psalm-142}

O Lord, hear my prayer, give ear unto my supplication in Thy truth; 
hearken unto me in Thy righteousness. And enter not into judgment with Thy servant, for in Thy sight shall no man living be justified.
 For the enemy hath persecuted my soul; 
he hath humbled my life down to the earth.
 He hath sat me in darkness as those that have been long dead, and my spirit within me is become despondent; 
within me my heart is troubled.
 I remembered days of old, I meditated on all Thy works, I pondered on the creations of Thy hands. 
I stretched forth my hands unto Thee; my soul thirsteth after Thee like a waterless land.
 Quickly hear me, O Lord; 
my spirit hath fainted away.
 Turn not Thy face away from me, lest I be like unto them that go down into the pit.
 Cause me to hear Thy mercy in the morning; 
for in Thee have I put my hope.
 Cause me to know, O Lord, the way wherein I should walk; 
for unto Thee have I lifted up my soul.
 Rescue me from mine enemies, O Lord; 
unto Thee have I fled for refuge. 
Teach me to do Thy will, for Thou art my God.
 Thy good Spirit shall lead me in the land of uprightness; 
for Thy name's sake, O Lord, shalt Thou quicken me.
 In Thy righteousness shalt Thou bring my soul out of affliction, and in Thy mercy shalt Thou utterly destroy mine enemies.
 And Thou shalt cut off all them that afflict my soul, for I am Thy servant.

Glory to the Father, and to the Son, and to the Holy Spirit, both now and ever, and unto the ages of ages.~Amen.

Alleluia, alleluia, alleluia. Glory to Thee, O God.
 \emph{Thrice.}

\textbf{\emph{Choir:}}~\emph{(In the tone of the troparion to be
chanted)}
: God is the Lord, and hath appeared unto us. Blessed is he that cometh in the name of the Lord.

\textbf{\emph{Reader:}}

\textbf{\emph{Stichos 1:}}
 O give thanks unto the Lord, for He is good, for His mercy endureth forever.

\textbf{\emph{Choir:}}
 God is the Lord, and hath appeared unto us.
 Blessed is he that cometh in the name of the Lord.

\textbf{\emph{Stichos 2:}}
 Surrounding me they compassed me, and by the name of the Lord I warded them off.

\textbf{\emph{Choir:}}
 God is the Lord\ldots.

\textbf{\emph{Stichos 3:}}
 I shall not die, but live, and I shall tell of the works of the Lord.

\textbf{\emph{Choir:}}
 God is the Lord\ldots.

\textbf{\emph{Stichos 4:}}
 The stone which the builders rejected, the same is become the head of the corner.
 This is the Lord's doing, and it is marvelous in our eyes.

\textbf{\emph{Choir:}}
 God is the Lord\ldots.

\emph{\textbf{This Troparion,}
~\textbf{4th~Tone:}}

To the Theotokos let us run now most earnestly, we sinners all and wretched ones, and fall down in repentance calling from the depths of our souls:
 O Lady, come unto our aid, have compassion upon us; 
hasten thou for we are lost in a throng of transgressions.
 Turn not thy servants away with empty hands, for thee alone do we have as our only hope.
 \emph{Twice.}

Glory to the Father, and to the Son, and to the Holy Spirit, both now and ever, and unto the ages of ages.~Amen.

Never, O Theotokos, will we cease to speak of thy powers, unworthy as we are.
 For if thou didst not intercede in prayer, who would have delivered us from so many dangers? Who would have kept us free until now?
 Let us never forsake thee, O Lady, for thou dost ever save thy servants from all perils.

\textbf{Psalm 50}

Have mercy on me, O God, according to Thy great mercy; and according to the multitude of Thy compassions blot out my transgression. 
Wash me thoroughly from mine iniquity, and cleanse me from my sin.
~For I know mine iniquity, and my sin is ever before me.
~Against Thee only have I sinned and done this evil before Thee, that Thou mightest be justified
in Thy words, and prevail when Thou art judged. 
For behold, I was conceived in iniquities, and in sins did my mother bear me. 
For behold, Thou hast loved truth; the hidden and secret things of Thy wisdom hast Thou made manifest unto me. 
Thou shalt sprinkle me with hyssop, and I shall be made clean; 
Thou shalt wash me, and I shall be made whiter than snow. 
Thou shalt make me to hear joy and gladness; 
the bones that be humbled, they shall rejoice. 
Turn Thy face away from my sins, and blot out all mine iniquities. 
Create in me a clean heart, O God, and renew a right spirit within me. 
Cast me not away from Thy presence, and take not Thy Holy Spirit from me. 
Restore unto me the joy of Thy salvation, and with Thy governing Spirit establish me. 
I shall teach transgressors Thy ways, and the ungodly shall turn back unto Thee. 
Deliver me from blood-guiltiness, O God, Thou God of my salvation; 
my tongue shall rejoice in Thy righteousness. 
O Lord, Thou shalt open my lips, and my mouth shall declare Thy praise. 
For if Thou hadst desired sacrifice, I had given it; 
with whole-burnt offerings Thou shalt not be pleased. 
A sacrifice unto God is a broken spirit; 
a heart that is broken and humbled God will not despise. 
Do~good, O Lord, in Thy good pleasure unto Zion, and let the walls of Jerusalem be builded. 
Then shalt Thou be pleased with a sacrifice of righteousness, with oblation and whole-burnt offerings. 
Then shall they offer bullocks upon Thine altar.

\textbf{\emph{Senior~Reader:}}
 O Lord, Jesus Christ, Son of God, have mercy on us.

\textbf{\emph{Choir:}}
 Amen.
\subsection{Canon}

\ins{Now the canon.}

\textbf{Then.}

It is truly meet to bless thee, the Theotokos, ever-blessed and most blameless, and Mother of Our God.
 More honorable than the Cherubim, and beyond compare more glorious than the Seraphim, who without corruption gavest birth to God the Word, the very Theotokos, thee do we magnify.
\emph{(Prostration to the floor)}

\textbf{\emph{Then the prayer for the canon to the Theotokos:}}

O our most blessed Queen, O Theotokos our hope, guardian of orphans, intercessor for strangers,~joy~of the sorrowful, protectress of the wronged:
 thou seest our misfortune, thou seest our affliction; 
help us, for we are infirm; 
feed us, for we are strangers.
 Thou knowest our offence: 
absolve it as thou wilt, for we have~no other help beside Thee, no other intercessor, nor good consoler, except thee,~O Mother of God.
 Do thou preserve and protect us unto the ages of ages.

\textbf{\emph{Choir:}}
 Amen.

%\subsubsection{End of Canon}\label{end-of-canon}

\textbf{\emph{Then:}}

Holy God, Holy Mighty, Holy Immortal, have mercy on us.
~\textbf{\emph{(Thrice)}}

Glory to the Father, and to the Son, and to the Holy Spirit, both now and ever, and unto the ages of ages.~Amen.

O Most Holy Trinity, have mercy on us. 
O Lord, blot out our sins. 
O Master, pardon our iniquities. 
O Holy One, visit and heal our infirmities for Thy name's sake.

Lord, have mercy.
~\textbf{\emph{(Thrice)}}

Glory to the Father and to the Son and to the Holy Spirit, both now and ever, and unto the ages of ages.~Amen.

Our Father, Who art in the Heavens, hallowed be Thy Name. 
Thy Kingdom come, Thy will be done, on earth as it is in Heaven. 
Give us this day our daily bread, and forgive us our debts, as we forgive our debtors;
and lead us not into temptation, but deliver us from the evil one.

\textbf{\emph{Senior Reader:}}
 O Lord, Jesus Christ, Son of God, have mercy on us.

\textbf{\emph{Reader:}}
~Amen.

\textbf{Troparion to St Gregory Palamas Tone 8:}

O luminary of Orthodoxy, support and teacher of the Church, ideal of monks and invincible champion of theologians, O wonderworker Gregory, boast of Thessalonika and herald of grace, always intercede for all of us that our souls may be saved.

Glory to the Father, and to the Son, and to the Holy Spirit:

\textbf{Kontakion Tone 8:}

\emph{(Podoben: ``O Victorious Leader\ldots{}'')}

Holy and divine instrument of wisdom, radiant and harmonious trumpet of theology, we praise thee in song, O divinely-speaking Gregory. 
As a mind standing before the Primal Mind, guide our minds to Him, Father, so that
we may cry aloud to thee: 
``Rejoice, herald of grace.''

Now and ever, and to the ages of ages. Amen.

Rejoice! O Virgin Theotokos! 
Mary, full of Grace, the Lord is with thee.
Blessed art thou among women, and blessed is the fruit of thy womb, for thou hast borne the Savior of our souls!

\textbf{\emph{Choir:}}
 Lord, have mercy.
 \emph{40x.}

\textbf{\emph{Senior Reader:}}
 Glory to the Father, and to the Son, and to the Holy Spirit, both now and ever, and unto the ages of ages. 
Amen.

\textbf{\emph{Choir:}}
~More honorable than the Cherubim, and beyond compare more glorious than the Seraphim; 
who without corruption gavest birth to God the Word, the very Theotokos, thee do we magnify.

Glory to the Father, and to the Son, and to the Holy Spirit, both now and ever, and unto the ages of ages.
~Amen.

Lord, have mercy.
 (Thrice)

O Lord, Bless.

\textbf{\emph{Senior Reader (Facing the East, rather than facing the
people):}}

O Lord Jesus Christ, Son of God, for the sake of the prayers of Thy most pure Mother, of our holy and God-bearing fathers, of Saint Gregory Palamas, Archbishop of Thessalonica, of \_\_\_\_\_ (saints of the day), and all the saints, have mercy on us and save us, for Thou art good and the Lover of mankind.

\textbf{\emph{Choir:}}
 Amen.

\textbf{Prayer for the Palamites:}

Lord Jesus Christ, save and protect those who have committed to praying this moleben with us daily. 
May this commitment bring us closer to Thee and by coming closer to Thee closer to one another and all of creation.
Through the prayers of Our Lady, The Theotokos and Ever Virgin Mary and St Gregory Palamas and all the saints. Amen.

